# Second break-out room  

We will calculated four different beta diversities: Bray-Curtis, Jaccard, unweighted UniFrac and Weighted UniFrac and use Principal Coordinates Analysis (PCoA, = Multidimensional scaling, MDS). This is a method to explore and to visualize similarities or dissimilarities of data.  

## Calculate beta-diversities

```{r, warning=FALSE, message=FALSE}
set.seed(12345)
metadata <- as(sample_data(ps1), "data.frame")

ord <- ordinate(ps1, method = "PCoA", distance = "bray") 
p <- plot_ordination(ps1, ord, color = "Feed",  label="Pen")  #shape= "Pen",
p1 <- p + theme_bw()  + ggtitle("PCoA Jaccard") 
p1

```

```{r, warning=FALSE, message=FALSE}
ord <- ordinate(ps1, method = "PCoA", distance = "jaccard", binary = TRUE) ##binary = TRUE!
p <- plot_ordination(ps1, ord, color = "Feed",  label="Pen")  #shape= "Pen",
p1 <- p + theme_bw()  + ggtitle("PCoA Jaccard") 
p1

```

```{r, warning=FALSE, message=FALSE}

ord <- ordinate(ps1, method = "PCoA", distance = "wunifrac") 
p <- plot_ordination(ps1, ord, color = "Feed",  label="Pen")  #shape= "Pen",
p1 <- p + theme_bw()  + ggtitle("PCoA wunifrac") 
p1
```

```{r, warning=FALSE, message=FALSE}
ord <- ordinate(ps1, method = "PCoA", distance = "unifrac") 
p <- plot_ordination(ps1, ord, color = "Feed", label="Pen")  #shape= "Pen",
p1 <- p + theme_bw()  + ggtitle("PCoA unifrac") 
p1
```

*What stands out?*   


*Which beta diversity would you use?*   


## PERMANOVA
```{r}
dist.b <- phyloseq::distance(ps1, method = "bray")
dist.j <- phyloseq::distance(ps1, method = "jaccard", binary = TRUE)
dist.uf <- phyloseq::distance(ps1, method = "unifrac")
dist.wuf <- phyloseq::distance(ps1, method = "wunifrac")


adonis(dist.b ~ Feed, data = metadata, perm=9999) 
ps.disper <- betadisper(dist.b, meta$Feed)
permutest(ps.disper)

adonis(dist.j ~ Feed, data = metadata, perm=9999) 
ps.disper <- betadisper(dist.j, meta$Feed)
permutest(ps.disper)

adonis(dist.uf ~ Feed, data = metadata, perm=9999) 
ps.disper <- betadisper(dist.uf, meta$Feed)
permutest(ps.disper)

adonis(dist.wuf ~ Feed, data = metadata, perm=9999) 
ps.disper <- betadisper(dist.wuf, meta$Feed)
permutest(ps.disper)

```


*Which beta diversity explained most of the variation with the factor feed?* 


> Additional reading
> [Gloor et al., 2017: Microbiome Datasets Are Compositional: And This Is Not Optional](https://www.frontiersin.org/articles/10.3389/fmicb.2017.02224/full) 
> [Paliy and Shankar, 2016: Application of multivariate statistical techniques in microbial ecology](https://doi.org/10.1111/mec.13536)
